
title: "IAM557_Term Project"
author: "Nika Rasoolzadeh"
date: "24/01/2021"
output: html_document
---
•
## Reading the Data :

In this project the \textbf{diabetes} dataset from LARS library is used. 

```{r}
                                   
library(lars);
data(diabetes);
patient_data <- data.frame(cbind(diabetes$x, y = diabetes$y))
a <- data.frame(cbind(diabetes$x2, y = diabetes$y))
n = dim(patient_data)[1]

```

## Creating Train and Test sets:

80%/20% random split of the original data

```{r}
set.seed(44) 
sample_test <- sample(n, round(n/5)) 

data.test <- patient_data[sample_test,]
data.train <- patient_data[-sample_test,]

```

## Define Predictor Matrix and Response Vector
```{r}
x <- model.matrix(y ~ ., data = patient_data)[,-1]

x.test <- x[sample_test,] 
x.train <- x[-sample_test,]
y <- patient_data$y
y.test <- y[sample_test]
y.train <- y[-sample_test]

# Getting the train and test set sizes:
train.size <- dim(data.train)[1] 
test.size <- dim(data.test)[1] 
```

## Linear Regression:

```{r}

lm.fit <- lm(y ~ ., data = data.train) 
summary(lm.fit)

```

```{r}
par(mfrow = c(2, 2)) 
plot(lm.fit) 
```

```{r}
names(lm.fit)
coef(lm.fit)
```

### Predict on test set:

```{r}
test.predict = predict(lm.fit, data.test)
test_MSE <- mean((y.test - test.predict)^2)
test_SE <- sd((y.test - test.predict)^2)/sqrt(test.size)
test_MSE; test_SE
```

## Best Subsest Selection
### Forward Stepwise:

```{r}
library(leaps)
set.seed(22)
regfit.fwd <- regsubsets(y ~ . , data = data.train, nvmax = 10,  method = "forward")
regfit.fwd.summary <- summary(regfit.fwd)
regfit.fwd.summary
```

### Forward Selection Plots

```{r}
par(mfrow = c(2, 2))
plot(regfit.fwd.summary$cp, xlab = "Number of variables", ylab = "C_p", type = "l")
points(which.min(regfit.fwd.summary$cp), regfit.fwd.summary$cp[which.min(regfit.fwd.summary$cp)],
       col = "red", cex = 2, pch = 20)

plot(regfit.fwd.summary$bic, xlab = "Number of variables", ylab = "BIC", type = "l")
points(which.min(regfit.fwd.summary$bic), regfit.fwd.summary$bic[which.min(regfit.fwd.summary$bic)], col = "red", cex = 2, pch = 20)

plot(regfit.fwd.summary$adjr2, xlab = "Number of variables", ylab = "Adjusted R^2", type = "l")
points(which.max(regfit.fwd.summary$adjr2), regfit.fwd.summary$adjr2[which.max(regfit.fwd.summary$adjr2)], col = "red", cex = 2, pch = 20)

mtext("Plots of Cp, BIC and adjusted R^2 for forward stepwise selection", side = 3, line = -2, outer = TRUE)
```

```{r}
coef(regfit.fwd, which.min(regfit.fwd.summary$bic))
```

### Inference:

```{r}
# The function below is defined for regsubsets since it does not have a 
# built in predict functin

predict.regsubsets <- function(object, newdata, id,...){
  form <- as.formula(object$call[[2]])
  mat <- model.matrix(form, newdata)
  cof <- coef(object, id = id)
  xvars <- names(cof)
  mat[, xvars]%*%cof
}

## 6 Parametered Model
regfit.fwd.pred = predict(regfit.fwd, data.test, id = 6)
fwd_test_MSE <- mean((y.test - regfit.fwd.pred)^2)
fwd_test_SE <- sd((y.test - regfit.fwd.pred)^2)/sqrt(test.size)

## 8 Parameter Model
regfit.fwd.pred2 = predict(regfit.fwd, data.test, id = 8)
fwd_test_MSE2 <- mean((y.test - regfit.fwd.pred2)^2)
fwd_test_SE2 <- sd((y.test - regfit.fwd.pred2)^2)/sqrt(test.size)

sprintf("6 predictor model test MSE: %0.3f, SE:%0.3f, 8 predictor model test MSE:%0.3f, SE:%0.3f",fwd_test_MSE,
        fwd_test_SE, fwd_test_MSE2, fwd_test_SE2)
```

### Backward Stepwise:

```{r}

set.seed(33)
regfit.bwd <- regsubsets(y ~ . , data = data.train, nvmax = 10,  method = "backward")
regfit.bwd.summary <- summary(regfit.bwd)
regfit.bwd.summary
```


### Backward Selection Plots

```{r}
par(mfrow = c(2, 2))
plot(regfit.bwd.summary$cp, xlab = "Number of variables", ylab = "C_p", type = "l")
points(which.min(regfit.bwd.summary$cp), regfit.bwd.summary$cp[which.min(regfit.bwd.summary$cp)],
       col = "red", cex = 2, pch = 20)

plot(regfit.bwd.summary$bic, xlab = "Number of variables", ylab = "BIC", type = "l")
points(which.min(regfit.bwd.summary$bic), regfit.bwd.summary$bic[which.min(regfit.bwd.summary$bic)], col = "red", cex = 2, pch = 20)

plot(regfit.bwd.summary$adjr2, xlab = "Number of variables", ylab = "Adjusted R^2", type = "l")
points(which.max(regfit.bwd.summary$adjr2), regfit.bwd.summary$adjr2[which.max(regfit.bwd.summary$adjr2)], col = "red", cex = 2, pch = 20)

mtext("Plots of Cp, BIC and adjusted R^2 for backward stepwise selection", side = 3, line = -2, outer = TRUE)
```

```{r}
coef(regfit.fwd, which.min(regfit.bwd.summary$cp))
```

### Inference:

```{r}

## 6 Parametered Model
regfit.bwd.pred = predict(regfit.bwd, data.test, id = 6)
bwd_test_MSE <- mean((y.test - regfit.bwd.pred)^2)
bwd_test_SE <- sd((y.test - regfit.bwd.pred)^2)/sqrt(test.size)

## 8 Parameter Model
regfit.bwd.pred2 = predict(regfit.bwd, data.test, id = 8)
bwd_test_MSE2 <- mean((y.test - regfit.bwd.pred2)^2)
bwd_test_SE2 <- sd((y.test - regfit.bwd.pred2)^2)/sqrt(test.size)

sprintf("6 predictor model test MSE: %0.3f, SE:%0.3f, 8 predictor model test MSE:%0.3f, SE:%0.3f",bwd_test_MSE,
        bwd_test_SE, bwd_test_MSE2, bwd_test_SE2)
```


## Lasso Regression:

```{r}
library(glmnet)
set.seed(55)
cv.lasso <- cv.glmnet(x.train, y.train, alpha = 1, nfolds = 10)
plot(cv.lasso)

```

```{r}
best_lambda <- cv.lasso$lambda.min
best_lambda; log(best_lambda)
```

```{r}
lasso.reg <- glmnet(x.train, y.train, alpha = 1, lambda = best_lambda)
lasso.pred <- predict(lasso.reg, s = best_lambda, newx = x.test)
coef.glmnet(lasso.reg)[1:11, ]
```

```{r}
lasso.test_MSE <- mean((y.test - lasso.pred)^2)
lasso.test_SE <- sd((y.test - lasso.pred)^2)/sqrt(test.size)
lasso.MAE <- mean(abs(y.test - lasso.pred))
sprintf("Lasso Regression test MSE: %0.3f, SE:%0.3f, MAE:%0.3f", lasso.test_MSE, lasso.test_SE, lasso.MAE)
```

## Comparing All the Results

```{r}
allmodels = data.frame(Model = c("Linear Regression Model", "Forward Selection Model with 6 predictor", "Forward Selection Model with 8 predictor", "Backward Selection Model with 6 predictor", "Backward Selection Model with 8 predictor","Lasso Model with 10 fold CV"), "Test MSE" = c(test_MSE, fwd_test_MSE, fwd_test_MSE2, bwd_test_MSE, bwd_test_MSE2, lasso.test_MSE), "test SE" = c(test_SE, fwd_test_SE, fwd_test_SE2, bwd_test_SE, bwd_test_SE2, lasso.test_SE))
print(allmodels)
```

## Tree Based Method:

```{r}
library(tree)
set.seed(66)
tree.diabetes = tree(data.train$y ~. , data= data.train)
summary(tree.diabetes)
```


```{r}
plot(tree.diabetes)
text(tree.diabetes, pretty=0)
```

### Prediction:
```{r}
tree.pred <- predict(tree.diabetes, newdata = data.test)
plot(tree.pred, y.test)
abline(0,1)
```

```{r}
tree.MSE <- mean((tree.pred - y.test)^2)
tree.MSE
```

### Cross Validation 
```{r}
cv.diabetes = cv.tree(tree.diabetes)
plot(cv.diabetes$size, cv.diabetes$dev,type='b')
```
### Pruning
```{r}
prune.diabetes = prune.tree(tree.diabetes, best=6)
plot(prune.diabetes)
text(prune.diabetes, pretty=0)
```
```{r}
ptree.pred <- predict(prune.diabetes, newdata = data.test)
plot(ptree.pred, y.test)
abline(0,1)
```

```{r}
ptree.MSE <- mean((ptree.pred - y.test)^2)
ptree.MSE
```

## Bagging and Random Forests:

```{r}
library(randomForest)
set.seed(77)
bag.diabetes = randomForest(y~.,data = data.train, mtry=10,importance=TRUE)
bag.diabetes

```

```{r}
bag.pred = predict(bag.diabetes, newdata= data.test)
plot(bag.pred, y.test)
abline(0,1)
```

```{r}
bag.MSE <- mean((bag.pred - y.test)^2)
bag.MSE
```
```{r}
set.seed(1)
rf.diabetes = randomForest(data.train$y~., data = data.train, mtry=3,importance=TRUE)
rf.pred =  predict(rf.diabetes, newdata = data.test)
re.MSE = mean((rf.pred-y.test)^2)
re.MSE
```
```{r}
importance(rf.diabetes)
```

```{r}
varImpPlot(rf.diabetes)
```

## Classification:

```{r}
library(class)
library(ggplot2)
library(caret)
set.seed(557)
diab_data <- patient_data
diab_sex = ifelse(diab_data$sex > 0, 1,0)
diab_data$sex = as.numeric(diab_sex)

m = dim(diab_data)[1]
sampletest <- sample(m, round(m/5)) 

diab_test <- diab_data[sampletest,]
diab_train <- diab_data[-sampletest,]

glm.fits = glm(factor(sex) ~. , data= diab_train ,family="binomial")
summary(glm.fits)
```

```{r}

train_pred <- predict(glm.fits, type = "response")
train.classes <- ifelse(train_pred > 0.5 , 1, 0)
table(train.classes, diab_train$sex)
```

```{r}
confusionMatrix(table(train.classes, diab_train$sex))
```

```{r}

test_pred <- predict(glm.fits, newdata = diab_test, type = "response")
test.classes <- ifelse(test_pred > 0.5 , 1, 0)
table(test.classes, diab_test$sex)
```

```{r}
confusionMatrix(table(test.classes, diab_test$sex))
```
```{r}
coef(glm.fits)
```

## KNN

```{r}
set.seed(63)
knn.model <- knn3(factor(sex) ~. , data = diab_train, k=4)
knn.probs <- predict(knn.model, newdata=diab_train )
knn.classes <- as.numeric(knn.probs[,2] > 0.5)
confusionMatrix(table(knn.classes, factor(diab_train$sex)))
```

```{r}
knn.probs.test <- predict(knn.model, newdata=diab_test )
knn.classes.test <- as.numeric(knn.probs.test[,2] > 0.5)
confusionMatrix(table(knn.classes.test, factor(diab_test$sex)))
```


# Validation Set Approach 
```{r}
valid_rnos = c(2, 47 ,112 ,86 ,69 ,34, 23 ,400, 357, 96, 80, 44, 77)
test.valid = diab_data[valid_rnos,]
train.valid <- diab_data[-valid_rnos,]
model <- glm(train.valid$sex ~., data = train.valid, family = binomial)
pred <- predict(model, newdata = test.valid, type = "response")
classes <- ifelse(pred > 0.5, 1,0)

if (nrow(train.valid >1)){
  conf.matrixv <- confusionMatrix(table(classes, test.valid$sex))
  missclass_ratev <- as.numeric(1-conf.matrixv$overall["Accuracy"])
} else {
  conf.matrixv = "not defined for a single observation"
  missclass_ratev = as.numeric(classes != test.valid$sex)
}
conf.matrixv
```
## Quadratic Linear Discriminant Analysis

```{r}
set.seed(99)
qda.model <- qda(factor(sex) ~., data = diab_train)
qda.model
```

```{r}
predmodel.test.qda = predict(qda.model, newdata = diab_test)
confusionMatrix(table(Predicted= predmodel.test.qda$class, sex=diab_test$sex))
```

